Exploring probability distributions for bivariate temporal granularities

### 2019-11-17 main.Rmd
cyclic time change

@@

### Circular {#circular-gran-def}

In a cyclic organization of time, the domain is composed of a set of recurring time values. The formalism of time granularities in cyclic arrangement is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data. Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index.

A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years. Since time is arranged in a cyclic way typically to accomodate periodicity, we will try to formalize how periodic linear granularities can be represented in a cyclic way. Since the assumption in [@Bettini2000-vy] is that all granularities can be generated from the bottom granularity by calendar algebraic operations - we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time using modular arithmetic.

Suppose,
z denotes the index of the tsibble,
x,y are two units in the hierarchy table with  $order(x) <  order(y)$. 
Let $f(x, y)$ denotes the accessor function for representing the finer unit x in terms of a coarser unit y and $c(x, y)$ is a constant which relates x and y.

>>>>

### Cyclic {#circular-gran-def}

In a cyclic organization of time, the domain is composed of a set of recurring time values. Hence, any time value can be preceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also suceeds Wednesday). Intutively, we need to manipulate the indices of periodic linear granularities so that it leads to repetitive patterns. The formalism of cyclic time granularities is attempted through the tsibble [@Wang2019-aw] framework of organizing temporal data. Suppose we have a tsibble with a time index in one column and keys and variables in other columns, a time domain, as defined by Bettini, is essentially a mapping of the index set to the time index. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years.

<!-- Since time is arranged in a cyclic way typically to accomodate periodicity, we will try to formalize how periodic linear granularities can be represented in a cyclic way. -->

The assumption in [@Bettini2000-vy] is that all linear granularities can be generated from the bottom granularity by calendar algebraic operations. Hence we assume that a bottom granularity exists and is represented by the index of the tsibble.

Suppose,
z denotes the index of the tsibble. Also represents the bottom granularity.
Let G be a periodic linear granularity, not equal to the bottom granularity, with each granule of G made up of $const(z, G)$ many granules of the bottom granularity.
C be the cyclic representation of the linear time granularity in terms of the bottom granularity.

Then, $$C(z, G) = z \mod c(z,G)$$

Now, suppose we want a cyclic representation of two periodic granularities.

$$C(G, H) = \lfloor z/const(z,G) \rfloor\mod const(G,H)$$

C1 and C2 are two cyclic granularities in the hierarchy table with  $order(C1) <  order(C2)$. 
Let $f(C_1, C_2)$ denotes the accessor function for representing the finer unit C1 in terms of a coarser unit C2 and $const(C1, C2)$ is a constant which relates C1 and C2.

we extend this notion to relate index of the two granularities in a way, which represents the circular arrangement of time using modular arithmetic.

Extending from the idea of linear granulariites, where 
$H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$ 
can be re-written as,
$H (j) = \bigcup_{i \in S_j \mod R}G((P/R) *(j - j \mod R) + i)$

The index in the right hand side looks like an equivalence class with modulus P/R.  
The value of $j\mod R$ is j for j< R and 1, 2, ...  for j>R.

Here, i runs from 1 to no. of granules of G and j runs from 1 to number of granules of H.
Hence, $j*\lfloor P/R \rfloor + i$.

For j = 1, we have ($\lfloor P/R \rfloor + i$)
For j = 2,  we have ($2*\lfloor P/R \rfloor + i$)

$H (j) = \bigcup_{i \in S_j}G(P*\lfloor j/R \rfloor + i)$

@@
